---
title: "STT 511 HW 2"
author: "Thomas Billman"
date: "January 29, 2018"
output: html_document
---
#R code for solving problem 3.7
We begin by reading in the data and finding the means
```{r data reading}
tensile.strength<-read.table(file = url("https://raw.githubusercontent.com/tbillman/Ghosh-511/master/Data/tensilestrength.txt"),header=T)
grp.means <- with(tensile.strength, tapply(TS,MT,mean))
```

First, we convert each variable to factor
```{r factoring}
tensile.strength$MT <- as.factor(tensile.strength$MT)
tensile.strength$run <- as.factor(tensile.strength$run)
```

Next, we run the model and find the analysis of variance
```{r AOV construction}
tensile.strength.aov <-aov(TS~MT,tensile.strength)
```
Testing the hypothesis (based on overall F-test)
```{r summary}
summary(tensile.strength.aov)
```

Extracting the residuals
```{r residuals}
aov_residuals <- resid(object=tensile.strength.aov)
```

Estimating the treatment means
```{r mean estimation}
model.tables(tensile.strength.aov)
```

Package for Fisher LSD comparison
```{r package installation}
#install.packages("agricolae") install this if you don't have it
library("agricolae")
```
Test
```{r Single difference test}
output<-LSD.test(tensile.strength.aov, "MT")
output
```

Tukey's multiple pairwise comparison
```{r Multiple Difference Test}
output<-HSD.test(tensile.strength.aov, "MT")
output
##(d), (e), and (f)
opar <- par(mfrow=c(2,2),cex=.8)
plot(tensile.strength.aov)
par(opar)
```

Constructing normal probability plot with theoretical line
```{r qqplots}
qqnorm(aov_residuals)

qqline(aov_residuals) 
```

Specific residual vs predictor plot
```{r Other residual plots}
aov_predictor<-fitted(tensile.strength.aov)

plot(aov_predictor,aov_residuals)
```
Producing a scatterplot
```{r General Scatterplot}
plot(tensile.strength.aov)
```

##Work for problem 3.10
Read in the data
```{r}
cot = cbind(c(rep("A",5), rep("B", 5),
              rep("C",5), rep("D", 5), rep("E",5)),
            c(7,7,15,11,9,
              12,17,12,18,18,
              14,19,19,18,18,
              19,25,22,19,23,
              7,11,10,15,11)
)
colnames(cot) = c("group","reading")
cot = as.data.frame(cot)
```
Factoring
```{r}
cot$group <- as.factor(cot$group)
cot$reading <- as.numeric(cot$reading)
```

Next, we run the model and find the analysis of variance
```{r AOV2 construction}
cot.aov <-aov(reading~group,cot)
```
Testing the hypothesis (based on overall F-test)
```{r}
summary(cot.aov)
```

Extracting the residuals
```{r}
aov_cot_residuals <- resid(object=cot.aov)
```

Estimating the treatment means
```{r}
model.tables(cot.aov)
```

LSD Comparison Test
```{r}
output<-LSD.test(cot.aov, "group")
output
```

Tukey's multiple pairwise comparison
```{r}
output<-HSD.test(tensile.strength.aov, "MT")
output
##(d), (e), and (f)
opar <- par(mfrow=c(2,2),cex=.8)
plot(tensile.strength.aov)
par(opar)
```

Constructing normal probability plot with theoretical line
```{r}
qqnorm(aov_residuals)

qqline(aov_residuals) 
```

Specific residual vs predictor plot
```{r}
aov_predictor<-fitted(tensile.strength.aov)

plot(aov_predictor,aov_residuals)
```
Producing a scatterplot
```{r}
plot(tensile.strength.aov)
```

##Work for problem 3.12
Read in the data
```{r}
dose = cbind(c(rep("A",4), rep("B", 4),
              rep("C",4)),
            c(24, 28, 37, 30,
              37, 44, 31, 35,
              42, 47, 52, 38)
)
colnames(dose) = c("dosage","activty")
dose = as.data.frame(dose)
```
Factoring
```{r}
dose$dosage <- as.factor(dose$dosage)
dose$activity <- as.numeric(levels(dose$activty)[as.numeric(dose$activty)])
```

Next, we run the model and find the analysis of variance
```{r}
dose.aov <-aov(activity~dosage,dose)
```
Testing the hypothesis (based on overall F-test)
```{r}
summary(dose.aov)
```

Extracting the residuals
```{r}
aov_dose_residuals <- resid(object=dose.aov)
```

Estimating the treatment means
```{r}
model.tables(dose.aov)
```

LSD Comparison Test
```{r}
output<-LSD.test(dose.aov, "dosage")
output
```

Tukey's multiple pairwise comparison
```{r}
output<-HSD.test(tensile.strength.aov, "MT")
output
##(d), (e), and (f)
opar <- par(mfrow=c(2,2),cex=.8)
plot(tensile.strength.aov)
par(opar)
```

Constructing normal probability plot with theoretical line
```{r}
qqnorm(aov_residuals)

qqline(aov_residuals) 
```

Specific residual vs predictor plot
```{r}
aov_predictor<-fitted(tensile.strength.aov)

plot(aov_predictor,aov_residuals)
```
Producing a scatterplot
```{r}
plot(tensile.strength.aov)
```


##Work for problem 3.27
Read in the data
```{r}
catl = cbind(c(rep("A",5), rep("B", 4),
              rep("C",3), rep("D",4)),
            c(58.2, 57.2, 58.4, 55.8, 54.9,
              56.3, 54.5, 57.0, 55.3,
              50.1, 54.2, 55.4,
              52.9, 49.9, 50.0, 51.7)
)
colnames(catl) = c("catalyst","concentration")
catl = as.data.frame(catl)
```
Factoring
```{r}
catl$catalyst <- as.factor(catl$catalyst)
catl$concentration <- as.numeric(levels(catl$concentration)[as.numeric(catl$concentration)])
```

Next, we run the model and find the analysis of variance
```{r}
catl.aov <-aov(concentration~catalyst,catl)
```
Testing the hypothesis (based on overall F-test)
```{r}
summary(catl.aov)
```

Extracting the residuals
```{r}
aov_catl_residuals <- resid(object=catl.aov)
```

Constructing normal probability plot with theoretical line
```{r}
qqnorm(aov_catl_residuals)

qqline(aov_catl_residuals) 
```

Specific residual vs predictor plot
```{r}
aov_predictor<-fitted(tensile.strength.aov)

plot(aov_predictor,aov_catl_residuals)
```
Confidence Intervals
```{r}
confint(catl.aov, level = .99)[1,]
```
