---
title: "STT 511 HW 2"
author: "Thomas Billman"
date: "January 29, 2018"
output: html_document
---
#R code for solving problem 3.7
We begin by reading in the data and finding the means
```{r data reading}
tensile.strength<-read.table(file = url("https://raw.githubusercontent.com/tbillman/Ghosh-511/master/Data/tensilestrength.txt"),header=T)
grp.means <- with(tensile.strength, tapply(TS,MT,mean))
```

First, we convert each variable to factor
```{r factoring}
tensile.strength$MT <- as.factor(tensile.strength$MT)
tensile.strength$run <- as.factor(tensile.strength$run)
```

Next, we run the model and find the analysis of variance
```{r AOV construction}
tensile.strength.aov <-aov(TS~MT,tensile.strength)
```
Testing the hypothesis (based on overall F-test)
```{r summary}
summary(tensile.strength.aov)
```

Extracting the residuals
```{r residuals}
aov_residuals <- resid(object=tensile.strength.aov)
```

Estimating the treatment means
```{r mean estimation}
model.tables(tensile.strength.aov)
```

Package for Fisher LSD comparison
```{r package installation}
#install.packages("agricolae") install this if you don't have it
library("agricolae")
```
Test
```{r Single difference test}
output<-LSD.test(tensile.strength.aov, "MT")
output
```

Tukey's multiple pairwise comparison
```{r Multiple Difference Test}
output<-HSD.test(tensile.strength.aov, "MT")
output
##(d), (e), and (f)
opar <- par(mfrow=c(2,2),cex=.8)
plot(tensile.strength.aov)
par(opar)
```

Constructing normal probability plot with theoretical line
```{r qqplots}
qqnorm(aov_residuals)

qqline(aov_residuals) 
```

Specific residual vs predictor plot
```{r Other residual plots}
aov_predictor<-fitted(tensile.strength.aov)

plot(aov_predictor,aov_residuals)
```
Producing a scatterplot
```{r General Scatterplot}
plot(tensile.strength.aov)
```