---
title: "STT511_HW5"
author: "Thomas Billman"
date: "April 3, 2018"
output: html_document
---
#Problem 1
a)
```{r}
glycogen <- c(106,101,120,86,123,97,
              58, 98, 85, 50, 111, 72,
              103, 84, 100, 88, 110, 91,
              53, 66, 61, 72, 85, 60)
trt <- c(rep("A",6),
         rep("a",6),
         rep("B",6),
         rep("b",6))
rat <- gl(6,1,24)
df <- data.frame(trt,rat,glycogen)
mod <- lm(glycogen ~., data = df)
anv1 <- anova(mod)
anv1
hsd <-TukeyHSD(aov(mod))
```
The following are the significant pairwise comparisons:
```{r}
which(hsd$trt[,4] < .05)
which(hsd$rat[,4] < .05)
```
b)
Are the contrasts orthogonal?
```{r}
a <- c(1,1,-1,-1)
b <- c(1,-1,1,-1)
c <- c(1,-1,-1,1)
sum(a*b)
sum(a*c)
sum(b*c)
```
Since the sum of all of these products are zero, all of these contrasts are orthogonal to one another.

c)
```{r}
tm <-sapply(c("A","a","B","b"), function(x){
  mean(df$glycogen[which(df$trt == x)])
})
n<- 6
ssdf <- apply(rbind(a,b,c),1, function(x){
  (sum(x*tm)^2)/((1/n)*sum(x^2))
})
mse <- anv1[3,3]
1-pf(ssdf/mse,1,24-4)
```
Since the P values for a and b are less than our $\alpha$ of .05, we reject $H_0$ and $H_1$  

#Problem 2 
This problem gives us the following information:  
```{r}
sst <- 1850
sstr <- 500
```
From here we can imply that 
```{r}
sse <- sst - sstr
```

```{r}
N <-30
a<-6
mstr <- sstr/(a-1)
mse <- sse/(N-a)
fval <- mstr/mse
pf(fval,a-1, N-a, lower.tail = F)
```
Given that our pval is greater than $\alpha$ of .05, we can not confirm that there exists at least one difference between the means.  

#Problem 3
a)
```{r}
tm <- c(3,1,0)
a <- c(-2,3,-1)
b <- c(1,0,-1)
c <- c(1,1,1)
d <- c(3,-2,1)
sum(a); sum(b); sum(c); sum(d)
sca <- sum(a*tm)
scb <- sum(b*tm)
n<- 9
mse <- 5
sea <-sqrt(mse * sum(a*a/n))
ta <- sca/sea
2*pt(ta,27-3)
```
Given that this value is above our $\alpha$ of .05, we fail to reject this hypothesis. 

b)

our t value will be double the first one
```{r}
seb <- sqrt(mse * sum(b*b/n))
tb <- scb/seb
2*pt(-tb, 27-3)
```
Given that this value is below our $\alpha$ of .05, we reject this hypothesis. 

c)
```{r}
scc <- sum(c*tm)
sec <- sqrt(mse * sum(c*c/n))
tc <- scc/sec
2*pt(-tc, 27-3)
```
Given that this value is below our $\alpha$ of .05, we reject this hypothesis. 

d)
```{r}
scd <- sum(d*tm)
sed <- sqrt(mse * sum(d*d/n))
td <- scd/sed
2*pt(-td, 27-3)
```
Given that this value is below our $\alpha$ of .05, we reject this hypothesis. 

ii)
To be a contrast, the sum of the coefficients on both sides of the equal sign must be equal. This is only true for a) and b)

iii)
```{r}
tm <- c(3,1,0)
a <- c(-2,3,-1)
b <- c(1,0,-1)
c <- c(1,1,1)
d <- c(3,-2,1)
sum(a); sum(b); sum(c); sum(d)
sca <- sum(a*tm)
scb <- sum(b*tm)
n<- 9
mse <- 5
sea <-sqrt(mse * sum(a*a/n))
sum(a*b)
```
given that a and b are the only contrasts and they are not orthogonal, there are not any orthogonal contrasts in this set of linear combinations.

iv)
This is not the same test. Even though we observed $\mu_3$ to be 0, there is still variability there that needs to be accounted for. The t test statistic for the testing the two hypothesis would need to be different, therefore they are not the same test.

#Problem 4
Knowing that:  
$Q_1 = 3\tau_1 - \tau_2 - \tau_3 - \tau_4$
$Q_2 = 2\tau_2 - \tau_3 - \tau_4$
$Q_3 = \tau_3 - \tau_4$
$SSTr = \sum_{i = 1}^{4}(\bar{y}_{i.}-\bar{y}_{..})^2$
$= \sum_{i = 1}^{4}\bar{y}_{i.}^2 - 4\bar{y}_{..}^2$
???

