---
title: "ICA4"
author: "Thomas Billman"
date: "March 21, 2018"
output: html_document
---
#Problem 1
###Reading in the data and performing an initial AOV
```{r}
x<- unlist(read.table(file=url("https://raw.githubusercontent.com/tbillman/Ghosh-511/master/Data/fertilizer.txt"), header = F))
miss_data<- data.frame(trt=gl(5,1,25),block=gl(5,5,25),x)
miss_data.aov <- aov(x~block+trt,miss_data)
miss_data.aov
```
###Part a) Write a linear model for this experiment making sure to explain all terms. You may assume that blocks are fixed.

For this experiment, given that there are treatments and blocks, the model would look like $y_{ij} = \mu + \tau_i + \beta_j + \epsilon_{ij}$ with $i \in 1...5$,
$j \in 1...5$, and $\epsilon_{ij}$ ~ $N(0,\sigma^2) \forall i,j$. We also assume all $\epsilon_{ij}$ are independent from one another.  

###Part b) Notice that in Block 1 there is a missing observation. Use the technique from the RCBD missing plot technique to impute this missing value. Then, compute the appropriate ANOVA table. Test for the equality of treatments. 

```{r}
newdat<- miss_data[which(!is.na(x)),]
r<- length(levels(newdat$trt))
b<- length(levels(newdat$block))
yoo <- sum(newdat$x)
i <- miss_data[which(is.na(x)),]$trt
j <- miss_data[which(is.na(x)),]$block
yio <- sum(newdat[which(newdat$trt == i ),]$x)
yoj <- sum(newdat[which(newdat$block == j),]$x)
a <- (r*yoj + b*yio - yoo)/((b-1)*(r-1))
```
Therefore we impute our x value as "a" or `r a`. We now sub this into the original data and rerun our ANOVA.
```{r}
idat <- miss_data
idat$x[3] <- a
lmi <- lm(x ~ block + trt, idat)
aovi <- anova(lmi)
aovi
```
This shows that since the P value for treatments is `r aovi[2,4]`, which is less than our $\alpha$ of .05. We conclude that there exists at least one difference between the treatment means and reject $H_o: \tau_1 = \tau_2 = \tau_3 = \tau_4 = \tau_5$.

###Part c) Provide all the corrections that need to be considered in theis case regarding the SS(Treatment) and SS(Block) as well as degrees of freedom. 

```{r}
csst <- (yoj - (r-1)*a)^2/(r*(r-1));csst
cssb <- (yio - (b-1)*a)^2/(b*(b-1));cssb
naovi <- aovi
naovi[3,1] <- 15
naovi[1,2] <- naovi[1,2] + cssb
naovi[2,2] <- naovi[2,2] + csst
naovi[3,2] <- sum(aovi$`Sum Sq`) - sum(naovi[1:2,2])
naovi[,3] <- naovi[,2]/naovi[,1]
for (q in 1:2){
  naovi[q,4] <- naovi[q,3]/naovi[3,3]
  naovi[q,5] <- pf(naovi[q,4], naovi[q,1], sum(naovi[,1]), lower.tail = F)
}
naovi
```

###Part d)
Add in a paragraph after confirming with Ghosh that this is how the corrections work.

