---
title: "ICA Orthogonal Contrasts"
author: "Thomas Billman"
date: "March 16, 2018"
output: html_document
---
#Problem 1
###Reading in Data
```{r Reading in Data}
dat <- c(2.04
, 1.91
, 2.00
, 1.92
, 1.85
, 1.96
, 1.88
, 1.90
, 1.66
, 1.71
, 1.42
, 1.76
, 1.66
, 1.61
, 1.55
, 1.66
, 1.53
, 1.54
, 1.38
, 1.31
, 1.35
, 1.27
, 1.26
, 1.38
, 1.15
, 1.22
, 1.17
, 1.16
, 1.21
, 1.28
, 1.17
, 1.26
, 0.83
, 1.08
, 1.02
, 1.09
, 1.06)
lev <- factor(c(gl(3,8),rep(4,7), rep(5,6)))
```

We also have to source our orthogonal contrast functions
```{r Source}
source("http://people.stat.sc.edu/hansont/stat506/cfcdae.R") 
```

###Initial Analysis
```{r}
lm1 <- lm(dat~lev)
anova(lm1)
```

###Contrasts
```{r}
A <- matrix(c(1,-1,0,0,0,
              0,0,1,-1,0,
              0,1,-1,0,0), ncol = 3, byrow = F)
A
linear.contrast(lm1,lev,A, jointF = T)
```

Given that at least one of the comparisons between means one through four are significant, we reject that the null hypothesis of $\mu_1 = \mu_2 = \mu_3 = \mu_4$. This is also true of the corresponding joint F test between our different hypotheses we constructed as the columns of our A matrix.

```{r}
linear.contrast(lm1,lev, c(.25,.25,.25,.25,-1))
```

Given that the P value for this linear contrast is well below .05, we reject our null hypothesis of : $H_0: \frac{\mu_1 + \mu_2 + \mu_3 + \mu_4}{4} - \mu_5 = 0$

```{r}
B <- matrix(c(1,-1,0,0,0,
              0,0,1,-1,0,
              0,1,-1,0,0,
              .25,.25,.25,.25,-1), ncol = 4, byrow = F)
B
linear.contrast(lm1,lev,B, jointF = T)
```
This analysis demonstrates the joint F value for the two contrasts. Since, the P value is well below .05, we determine that at least one of them needs to be rejected.

#Problem 2
###Question:
Let us suppose we have 5 treatments of an experiment. Namely, $\tau_i$ for $i \in 1...5$, From this create at least two linear contrasts such that: (a) they are pairwise orthogonal; and (b) they are not orthogonal. Be sure to justify your choise with all the details.  
  
###Part (a)
One pair of linear contrasts that would be pairwise orthogonal would be $\tau_1 = \tau_2$ and $\tau_3 = \tau_4$, this would let $c_i = \{1,-1,0,0,0\}$ and $d_i = \{0,0,1,-1,0\}$. Therefore both $c_i$ and $d_i$ are orthogonal because, $\sum_{i=1}^5 c_i = \sum_{i=1}^5 d_i = 0$. We also conclude that they are pairwise orthogonal because $\sum_{i=1}^5 c_i d_i = 0+0+0+0+0 = 0$

###Part (b)
One pair of linear contrasts that would NOT be pairwise orthogonal would be $\tau_1 = \tau_2$ and $\tau_2 = \tau_3$, this would let $c_i = \{1,-1,0,0,0\}$ and $d_i = \{0,1,-1,0,0\}$. Therefore both $c_i$ and $d_i$ are orthogonal because, $\sum_{i=1}^5 c_i = \sum_{i=1}^5 d_i = 0$. However, we also conclude that they are NOT pairwise orthogonal because $\sum_{i=1}^5 c_i d_i = 0+-1+0+0+0 = -1\neq0$.

#Problem 3
###Question:
Suppose, we have the following data set regarding to an industrial process with 4 operators and their strength in coninuing their work (in some appropriate measuring environment):
```{r}
strength=c(7.3550, 7.5900, 7.7750, 6.2350, 7.0000, 6.7900, 7.6250, 5.7975, 8.1050, 8.3600, 8.0100, 8.2675)
oper=gl(4,3)
```
Then,  

<ul>  
<li> Conduct separate tests (using the idea on linear contrast) for the following two sets of null hypotheses: $H_0: \mu_2 = \mu_3 = \mu_4$ and $H_0: \frac{\mu_1 + \mu_2 + \mu_3}{3} - \mu_4 = 0$ at the 5% level of significance. In other words, test that the contrast is zero. Provide the estimates for each of the contrasts associated with those two null hypotheses. </li>
<li> Conduct a joint F-test.</li>
</ul>

###Answer:
We begin by constructing a linear model to find the treatment effects and means of the different operators.
```{r}
lm2 <- lm(strength ~ oper)
anova(lm2)
```

###Contrasts
We can break down $\mu_2 = \mu_3 = \mu_4$ as the two pairwise comparisons of $\mu_2 = \mu_3$ and $\mu_3 = \mu_4$. This allows us to construct a matrix for comparison between these two constrasts.
```{r}
C <- matrix(c(0,1,-1,0,
              0,0,1,-1), ncol = 2, byrow = F)
C
lc1 <- linear.contrast(lm2, oper, C, jointF = T)
lc1
```

Given that the p value for the F test was `r lc1$Ftest[1,4]` which is not less than our $\alpha$ of .05, we fail to reject the null hypothesis of $\mu_2 = \mu_3 = \mu_4$.  
  
###Other Hypothesis
For our second null hypothesis of $H_0: \frac{\mu_1 + \mu_2 + \mu_3}{3} - \mu_4 = 0$, we test the contrast with the following code:  
```{r}
lc2 <- linear.contrast(lm2, oper, c(rep(1/3,3),-1))
lc2
```
Given that the p value for the F test was `r lc2[1,4]` which is less than our $\alpha$ of .05, we reject the null hypothesis of $\frac{\mu_1 + \mu_2 + \mu_3}{3} - \mu_4 = 0$.  

####Joint F Test
To conduct the Joint F Test between these two null hypotheses, we check construct a set of contrasts from both tests. This is done as follows:
```{r}
D <- matrix(c(0,1,-1,0,
              0,0,1,-1,
              rep(1/3,3),-1), ncol = 3, byrow = F)
D
lc3 <- linear.contrast(lm2, oper, D, jointF = T)
lc3
```
Given that the p value for the F test was `r lc3$Ftest[1,4]` which is greater than our $\alpha$ of .05, we fail to reject the set of hypotheses.
